### 2017-04-01 ThesisText/background.tex
Adds introduction
Adds conclusions and future work

\chapter{Background}

@@

\section{Cross Validation}

@@

Testing a model with different objects then the ones used in training is very
important. This is because a model, while training, will learn the very
specific details of that data and may over fit. When over fitted, a model will
not work for general examples of the data for which it was trained and will
only ``remember'' training data instead.

@@

Educational Data Mining, \gls{edm}, refers to the act of applying Data Mining
techniques to educational data. Educational data refers to data which origin is
in the act of teaching and training, in the activities of professors and
students, results of evaluation, and so on. In this project, the origin of data
is the University of Évora's Moodle.

@@

One such example of~\gls{edm} Data Mining tasks is too predict student grades
based features from the students themselves, the courses they are undertaking,
and their activity in a~\gls{lms}. By training such models, professors can make
decisions about how courses are organized regarding students success. For
example, a certain number of activities in a given course may be beneficial to
it's success because it keeps students engaged, but any number of activities
more might not make a difference.~\cite{ind_001, ind_002, ind_005}

@@

In~\cite{ind_006}, networks are trained from Moodle logs. The data from the
logs are structured in a slightly different way then the previous studies. The
features of this data are shown in table~\ref{tab:ind_006_fields}.

>>>>

\chapter{Background}
\label{sec:back}

@@

\subsection{Cross Validation}
\label{sec:cross_val}

@@

Testing a model with different objects than the ones used in training is very
important. This is because a model, while training, will learn the very
specific details of that data and may over fit. When over fitted, a model will
not work for general examples of the data for which it was trained and will
only ``remember'' training data instead.

@@

\acrfull{edm}, refers to the act of applying Data Mining techniques to
educational data. Educational data refers to data which origin is in the act of
teaching and training, in the activities of professors and students, results of
evaluation, and so on. In this project, the origin of data is the University of
Évora's Moodle.

@@

One such example of an~\gls{edm} tasks is to predict student grades based
features from the students themselves, the courses they are undertaking, and
their activity in a~\gls{lms}. By training such models, professors can make
decisions about how courses are organized regarding students success. For
example, a certain number of activities in a given course may be beneficial to
it's success because it keeps students engaged, but any number of activities
more might not make a difference.~\cite{ind_001, ind_002, ind_005}

@@

In~\cite{ind_006}, networks are trained from Moodle logs. The data from the
logs are structured in a slightly different way than the previous studies. The
features of this data are shown in table~\ref{tab:ind_006_fields}.

### 2017-04-01 ThesisText/data.tex
Minor corrections on statistics
Adds introduction

\chapter{University of Évora's Moodle Data}

@@

To begin with, we explore what was initially provided in terms of data. The
repository is a spreadsheet book with 12 sheets. The sheets contain raw data
and some calculated fields.

In order to explore the data using the bash tools (cut, grep, etc) and tools
like Python with Pandas, they were exported to a CSV format. Some
transformations were applied to these files in other to clean useless data
points or make the data easier to process.

@@

Some sheets contained information that was the result of calculations from the
established data. Analysis like these were removed from the data for two
reasons. First, some analysis are not useful so them being in the data only
makes the data more complex. By removing such analysis we get a simpler data
model. Second, even if some analysis are in fact useful, they are still easily
reproduced by the Python tools, with the difference that they are only
reproduced when needed for specific experiments.

Some analysis were made and are described in section~\ref{set:stats}

@@

@@

Moodle logs contain the actual activities of Moodle users. Each log entry
states a user who performed a~\gls{crud} action in a course page. The actions
are aggregated per week. Table~\ref{tab:moodle_logs_features} shows the
features of the Moodle logs, and table~\ref{tab:moodle_logs_enum} shows the
values of the enumerated fields. There are a total of 25207 entries in the
Logs.

@@

    \caption
        [Moode Logs features]
        {Moode Logs features.}

@@

From the original data we make an initial pre processing in order to have a
consistent repository of datasets ready for further analysis. The pre-processed
datasets have consistent names and fields.

@@

The \texttt{Profile} dataset contains the information stated in
section~\ref{sec:profile}. The fields are shown in
table~\ref{tab:profile_fields}.

>>>>

\chapter{University of Évora's Moodle Data}
\label{sec:data}

@@

To begin with, we explore the data which was initially provided. The repository
is a spreadsheet book with 12 sheets. The sheets contain raw data and some
calculated fields.

In order to explore the data using the bash tools (cut, grep, etc) and tools
like Python with Pandas, they were exported to a CSV format. Some
transformations were applied to these files in order to clean useless data
points or make the data easier to process.

@@

Some sheets contained information that was the result of calculations from the
established data. Analysis like these were removed from the data for two
reasons. First, some analysis are not useful so them being in the data only
make it more complex. By removing such analysis we get a simpler data
model. Second, even if some analysis are in fact useful, they are still easily
reproduced by the Python tools, with the difference that they are only
reproduced when needed for specific experiments.

@@

There are a total of 331 entries in the Profile dataset.

@@

Moodle logs contain the actual activities of Moodle users. Each log entry
states a user who performed a~\gls{crud} action in a course page. The actions
are aggregated per week. Table~\ref{tab:moodle_logs_features} shows the
features of the Moodle logs, and table~\ref{tab:moodle_logs_enum} shows the
values of the enumerated fields. There are a total of 25206 entries in the
Logs.

@@

    \caption
        [Moodle Logs features]
        {Moodle Logs features.}

@@

From the original data we make an initial preprocessing in order to have a
consistent repository of datasets ready for further analysis. The preprocessed
datasets have consistent names and fields.

@@

The \texttt{Profile} dataset contains the information stated in
section~\ref{sec:profile}. The fields are shown in
table~\ref{tab:profile_features}.

### 2017-04-01 ThesisText/statistics.tex

To begin exploring the courses some counts are made. We count the number of
overall courses, the number of departments, and number of degrees.
Table~\ref{tab:courses_stats} shows these counts. We see 77 courses which are
part of 9 different departments and make up 11 degrees. While the number of
courses is much greater then some experiments mentioned in the background, it
is important to understand how many students does each of these courses have.

@@

Table~\ref{tab:moodle_users_stats} shows some stats regarding Moodle users.
Some of the explored projects listed in the background, like~\cite{ind_001},
have a number of students around 800 to 900 which is greater then the number of
students in this dataset, but they also have a lower number of courses, for
example~\cite{ind_001} has 7, where this dataset has 77.

@@

    \caption
        [Professor features]
        {Professor features.}

@@

We take the two courses with the most activity overall and plot a graph which x
axis is the weeks and y axis is the number of activities.
Figure~\ref{fig:stat_004_all} shows the graph for these courses. The number of
read activities and number of total activities is much greater then the number
of create, update, and delete activities. Because of this, we plot a graph
which only contains create, update, and delete activities for the same courses
as shown in figure~\ref{fig:stat_004_cud}.

@@

The number of total activities per week drops significantly. This is because
some courses are longer then others. For example, some courses may have their
last evaluation in the first weeks of the exam season, while others may still
have ongoing activity after exam season.

@@

    \caption
        [Statistics on the Results]
        {Statistics on the Results.}

>>>>

To begin exploring the courses, some counts are made. We count the number of
overall courses, the number of departments, and number of degrees.
Table~\ref{tab:courses_stats} shows these counts. We see 77 courses which are
part of 9 different departments and make up 11 degrees. While the number of
courses is much greater than some experiments mentioned in the background, it
is important to understand how many students does each of these courses have.

@@

Table~\ref{tab:moodle_users_stats} shows some counts regarding Moodle users.
Some of the explored projects listed in the background, like~\cite{ind_001},
have a number of students around 800 to 900 which is greater than the number of
students in this dataset, but they also have a lower number of courses, for
example~\cite{ind_001} has 7, where this dataset has 77.

@@

    \caption
        [Moodle users counts]
        {Moodle users counts.}

@@

We take the two courses with the most activity overall and plot a graph which x
axis is the weeks and y axis is the number of activities.
Figure~\ref{fig:stat_004_all} shows the graph for these courses. The number of
read activities and number of total activities is much greater than the number
of create, update, and delete activities. Because of this, we plot a graph
which only contains create, update, and delete activities for the same courses
as shown in figure~\ref{fig:stat_004_cud}.

@@

The number of total activities per week drops significantly. This is because
some courses are longer than others. For example, some courses may have their
last evaluation in the first weeks of the exam season, while others may still
have ongoing activity after exam season.

@@

    \caption
        [Counts on the Results]
        {Counts on the Results.}
